---
title: "Generating the Representative Sample"
author: "Brian Waismeyer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating the Representative Sample}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This document walks through the code and logic used to generate a random sample 
of video games - stratified by year - from our video game population.

There are a number of ways to approach stratified sampling (e.g., see 
[here](http://stattrek.com/survey-research/stratified-sampling.aspx) for a 
summary of a few). Our objective is to generate a sample that allows us to 
estimate how specific features of video games have changed year-by-year. 
Additionally, as we'll see below, starting in the 1980s the number of games 
released (that we have records for) is surprisingly conistent.

Our obective and the nature of the data lend themselves well to fixed-sample 
stratified sampling. This approach can be problematic if the strata vary wildly 
in size and/or the variability of their contents. However, given the size of our
later strata, proportionate sampling would not have a large impact on many of
our strata sizes.

## Prepare the Workspace

We start by loading supporting packages and setting desired global options.

```{r prepare_workspace}
# Load supporting packages.
library(dplyr)          # Data manipulation.
library(tidyr)          # Data manipulation.
library(stringr)        # String maniuplation.
library(ggplot2)        # Data visualization.
```

To keep the later portions of the document concise, we also define our custom
helper functions here. Functions are presented in the order they are called.

```{r define_helper_functions}
# Generate the index for a stratified random sample.
#
# This function handles our stratified random sampling. It differs from other
# such sampling functions in that it can tolerate strata smaller than the fixed
# sample size while still sampling without replacement.
#
# @param df The dataframe the function will build an index for.
# @param strat_col A string naming the column the function will use for
#   stratification.
# @param sample_size Integer indicating the size of the sample to draw from
#   each stratum.
#
# @returns
#   Returns an index as a vector that can be used to subset the target dataframe
#   to generate the stratified random sample.
sample_strata <- function(df, strat_col, sample_size) {
    # Insure the formatting of the dataframe.
    df <- as.data.frame(df)
    
    # Define the bucket for values we'll be indexing.
    sample_index <- c()
    
    # Capture the unique strata values.
    strata <- sort(unique(df[, strat_col]), 
                   # A subtle little move - we retain our NAs and put them at the
                   # end of our unique list.
                   na.last = TRUE)
    
    # Sample from the strata.
    for(stratum in strata) {
        # Observe which records are associated with the strata. This is
        # structured to handle NAs.
        stratum_index <- which(df[, strat_col] %in% stratum)
        
        # Test if sufficiently large to sample from.
        if(length(stratum_index) <= sample_size) {
            # If too small, simply capture the indices for all of the records.
            sample_index <- c(sample_index, stratum_index)
        } else {
            # Otherwise, sample and capture sample_size records without 
            # replacement.
            random_index <- sample(stratum_index, sample_size, replace = FALSE)
            
            sample_index <- c(sample_index, random_index)
        }
    }
    
    # Return the sampled records.
    return(sample_index)
}
```

Finally, we load the dataframe we will be sampling from.

```{r load_merge_dfs}
# Load the video game "population" dataframe.
load("./merge_final_df.Rds")
```

## Generating the Stratified Sample

As noted above, we will be aiming to sample a fixed number of games from our 
"game population" per release year. We can think of this sample from two 
perspectives: as a sample from each year or as a single stratified sample.

The fixed size we choose will represent our individual year sample size. The 
collective, stratified sample is our representation of the entire population. We
will choose our fixed strata sample size to product a population sample size of
about 2500 games (~ 2.5% of the total population).

Where there are fewer games in a stratum than our fixed sample size, we will
simply sample the entire stratum.

Our data includes games from `r length(unique(final_df$first_release_year))`
years. For a naive estimate, we simply divide our target size by our number of
strata: 2500 /`r length(unique(final_df$first_release_year))` =
`r 2500 / length(unique(final_df$first_release_year))`. Of course, a number of
our early strata will have fewer games than our naive estimate, so we bump that
up slightly so that we sample more from our larger strata.

```{r sample_generation}
# Set the random sample seed (so that people can track how the current sample
# in particular was developed).
set.seed(2016)

# Generate a sample index, fully sampling those strata which are smaller than 
# the target stratum sample size.
strat_sample_index <- sample_strata(final_df, "first_release_year", 60)

# Extract our represenative sample records.
strat_sample_df <- final_df[strat_sample_index, ]

# Inspect the sample structure.
str(strat_sample_df)

# Inspect the sample by release year.
strat_sample_df %>%
    group_by(first_release_year) %>%
    summarise(count = n()) %>%
    data.frame()
```

We see that our early release years often have less than our target stratum 
sample size. However, from 1977 onwards at least 60 games were released per year
(including those games which do not have a properly recorded release year).

## Inspecting the Sample for Duplicates

Although we took great pains to remove duplicates from our game population, the 
truth is that we likely missed a few. The sample records were sorted by game 
title and inspected manually for duplicate entries. Duplicates are removed below
and replaced with a new randomly drawn title. The observed duplicate rate is
then used to estimate the rate and number of duplicates in the game population.

```{r replacing_duplicates}
# A sorted version of the representative sample to use during manual duplicate
# inspection.
sdf <- strat_sample_df %>%
    arrange(title) %>%
    select(title, first_release_year) %>%
    data.frame()

# Example of alphabetical inspection.
# sdf[1:100, ]

# Code use to generate similar string sets.
# sdf_dup_candidates <- fuzzy_match_all(sdf$title, remove_matches = TRUE,
#                                       max_dist = .2)
# not_na <- sapply(sdf_dup_candidates, function(x) {
#     is.na(x[[1]])
# })
# sdf_dup_candidates[not_na]

# Entries identified as duplicates. The selection was based initially on entry 
# quality. If quality was identical, it was based on whichever had an earlier 
# index.

# NOTE: I spell out the titles here largely for reader convenience. I actually
# use the list names (the sdf index values) to do the matching in case there are
# special characters in the title (e.g., Shui Hu Zhuan has special characters
# that can't be produced easily on an English system).
duplicates <- list(
    "69" = "Air Raid!",     
    "164" = "Asteroid",     
    "462" = "CORE: Cybernetic Organism Recovery Expedition",
    "498" = "Cross Channel",
    "1053" = "Intercepter",
    "1634" = "Ping-Pong",    
    "1930" = "Shui Hu Zhuàn",
    "2014" = "Space War",
    "2430" = "TV School House 2",
    "2489" = "Videocart 1: Tic Tac Toe: Shooting Gallery: Doodle: Quadradoodle",
    "2490" = "Videocart 10: Maze",
    "2492" = "Videocart 2: Desert Fox: Shooting Gallery",
    "2496" = "Videocart 7",
    "2575" = "Worm!",
    "2078" = "StarFox Command",
    "2281" = "The Fairly OddParents: Breakin' Da Rules",
    "2615" = "YuYu Hakusho"
)

duplicate_names <- sdf$title[as.integer(names(duplicates))]

# Now we flag the duplicates in both our sample dataframe and the game 
# population dataframe.
strat_sample_df$is_duplicate <- ifelse(strat_sample_df$title %in% 
                                           duplicate_names,
                                       TRUE,
                                       FALSE)

final_df$is_duplicate <- ifelse(final_df$title %in% 
                                    duplicate_names,
                                TRUE,
                                FALSE)

# And we flag which games have already been sampled in game population 
# dataframe.
final_df$is_sampled <- ifelse(final_df$title %in% strat_sample_df$title,
                              TRUE,
                              FALSE)

# Next we calculate our observed duplicate rate and confidence intervals.
duplication_rate <- length(duplicate_names) / nrow(strat_sample_df)

# Assuming that the distribution for sampled duplicate error rates is normally
# distributed, we estimate the actual rate of error in the game population.

# Calculate the standard error for a proportion.
ste_duplications <- sqrt(
    (duplication_rate * (1 - duplication_rate)) / 
        nrow(strat_sample_df)
)

# Select our confidence interval and its associated Z.
conf_z <- 1.96      # For 95% conf interval.

dup_rate_conf_size <- conf_z * ste_duplications

# State the results.
sprintf("Estimated rate of duplicates is %s (+/- %s).",
        round(duplication_rate, 3), round(dup_rate_conf_size, 3))
sprintf("Estimated number of duplicates (out of %s) is %s (+/- %s).",
        nrow(final_df), 
        round(nrow(final_df) * duplication_rate),
        round(nrow(final_df) * dup_rate_conf_size))

# Remove current intermediate values.
rm(conf_z, dup_rate_conf_size, duplicate_names, duplication_rate, 
   duplicates)

# Next we calculate how many games for each stratum need to be replaced.
strat_counts <- strat_sample_df %>%
    group_by(first_release_year) %>%
    summarise(games_to_replace = sum(is_duplicate)) %>%
    filter(games_to_replace > 0)

# We randomly sample an appropriate number of new records from each stratum, 
# ignoring games that have already been sampled. If all records in a statum have
# already been sampled, we get an NA and accept that we'll be shrinking that
# stratum.
strat_sample_index <- unlist(sapply(1:nrow(strat_counts), function(x) {
    current_year <- strat_counts$first_release_year[x]
    sample_size <- strat_counts$games_to_replace[x]
    
    stratum_index <- which((final_df$first_release_year %in% current_year) &
                               (!final_df$is_sampled))
    
    if(length(stratum_index) == 0) {
        return(NA)
    }
    
    sample_index <- sample(stratum_index, sample_size, replace = FALSE)
    
    return(sample_index)
}))

# We do a manual review of the newly sampled items to see if any of them in turn
# need replacement. None were identified as duplicates in the current sample
replacement_df <- final_df[strat_sample_index, ]

replacement_df %>%
    select(title, first_release_year)

replacement_df$is_duplicate <- FALSE
strat_sample_df$is_sampled <- TRUE

# Add the replacements to the sample.
strat_sample_df <- rbind(strat_sample_df, replacement_df)

# Drop the duplicates and any junk entries.
strat_sample_df <- strat_sample_df %>%
    filter(!is_duplicate,
           !is.na(title))
```

```{r first_pass_assignments}
# Create a version of the dataframe for use during first pass inspections of
# game records.

# First we select a target platform.
strat_sample_df$target_platform  <- sapply(
    strat_sample_df$platform, function(current_value) {
        if(!grepl("----", current_value)) {
            return(current_value)
        }
        
        value_set <- unlist(str_split(current_value, "----"))
        
        current_value <- sample(value_set, 1)
        
        return(current_value)        
}, USE.NAMES = FALSE)

# Then we make a composite column for copy-pasting into the survey.
strat_sample_df <- strat_sample_df %>%
    mutate(copy_col = str_c(str_replace_na(title), 
                            str_replace_na(first_release_year),
                            str_replace_na(target_platform),
                            sep = "----"))

# Then we divvy the table up into four parts and give these to team members.
strat_sample_df$reviewer <- c("Brian", "Kameron", "Kate", "Sam")

# Finally, we reorganize the results slightly for reviewer convenience.
strat_sample_df <- strat_sample_df %>%
    select(copy_col, reviewer, everything())

# We save the results for distriubtion to the team.
write.csv(strat_sample_df, file = "rep_sample_for_review.csv")
```
